Application of a Cut-Cell Immersed Boundary Method for Wildfire Simulation over Complex Terrain

### 2020-12-09 Response_to_A_Maranghides.tex

I am specifically interested in the case of the Cogoleto Fire. I think the case presented here is \textbf{oversimplified and can potentially direct the reader to make certain assumptions about the model capabilities}. Even though this is a small part of the paper in many ways it is maybe the most important in terms of representing the model usefulness.

What is lacking here are:
%
\begin{itemize}

   \item Information about the scale of the event: how large was the fire from origin to furthest point, how long it took to reach the final perimeter. A scale on the figure 4 would be helpful.

   \hl{This info was obtained from the simulation domain. From origin to furthest point measured distance is about 1200 m. The fire run overnight and onto the next day.}

@@

   \item Wind context: was the single measurement consistent both in speed and direction throughout the event. How representative was this single point measurement, given the second closest weather station and how that data compared. was the station at a specific location (ravine or local acceparation zone). I can tell you that in CA, in the Camp Fire (as an example here) 5 -8 km can make all the difference in the world (100 km/h vs 10 km/h)

   \hl{From the information provided from CIMA : Winds recorded at nearby sensor station were found predominantly coming from the north with magnitudes that ranged from 20 m/sec during the night and diminished to 5 m/sec towards the morning hours.}

@@

   \item information on defensive actions. Defensive actions not only impact the final perimeter but can also impact the direction as firefighters typically try to direct the fire away from high value targets. This can be done with both ground and aerial resources. I do not know what was used and when.

   \hl{Information from https://www.thelocal.it/20190326/residents-evacuated-as-wildfires-take-hold-near-genoa : Up to 70 firefighters worked at the scene, with help from helicopters and fire hydrants. Suppression work was directed into stopping fire spread into the coastal town of Cogoleto.}

@@

   \item There is no time scale provided on the runs or event progression so it is very difficult to judge what is presented.

   \item There is no mention of spotting within the fire perimeter and also outside. Short, intermediate and far field spotting can play critical roles in the evolution of a wildfire and/or WUI incident and in my mind is one of the key issues that realistically limits the applicability of predictive tools on that scale. Since we cannot resolve where embers will result in ignitions and the location of the actual ignitions can have dramatic impact on how the event develops, not knowing limits the usefulness of these tools. A probabilistic approach does not improve the usefulness in my mind.

   \hl{Ema, private communication : There's been a lot of spotting. Added : Further, spot ignition was reported by first responders at the scene, a common fire spread mechanism that remains a major issue in wildfire modeling.}

@@

   \item Explicit listing of the model limitations of dealing with fuel breaks both natural and manmade. While I appreciate that if the wind was truly 100km/h fuel breaks may have had little impact on the overall fire spread, the effect on fuel breaks can still provide critical information to understanding how fire behaved and should be discussed.

    \hl{Kevin : For the particle and boundary fuel models of flame spread, fuel breaks are handled directly by removing combustible vegetation from the model. For the level set method, a fire break is a region declared as "non-combustible." The 13 Rothermel fuel types have been supplemented by "water", "ice", "pavement" and various other non-combustible regions that would be tiled onto our terrain map.}

@@

   \item I would also like to see some statement explicitly addressing the complexity of dealing with non-uniform fuels (not only vegetative but also WUI). Even using the level set approach one needs to be mindful of the potential change in fire spread as the fire migrates between different fuels.

   \hl{Non-uniform fuels : Adressed in the statement on the complexity of simulation real WUI fires (last paragraph of cogoleto section.)}

### 2020-12-09 Vanella_Article_Atmosphere_2020.tex
Update on Overleaf.

\usepackage[export]{adjustbox}

@@

Wildland fire modeling typically focuses on tracking the fire front, which can have a complex shape that moves and deforms based on local conditions like fuel loading, moisture content, wind, and terrain. The fire front requires a mathematical representation. The Lagrangian approach is based on a polygonal mesh of control markers that represents the front~\cite{Finney:FARSITE,Bova:IJWF2015}. On the other hand, an Eulerian representation allows for a fixed two-dimensional mesh on which the fire front is defined implicitly by a scalar field. This scalar field is called the level set function and numerical methods that compute the function are called Level Set Methods (LSM)~\cite{Sethian:1999,Osher:2006}. Some wildfire solvers that use LSM to track the fire front are described in references~\cite{coen_2013,Bova:IJWF2015,FDS_Users_Guide,LAUTENBERGER_2013}. A practical comparison of the two techniques as implemented in the Fire Dynamics Simulator (FDS)~\cite{Mell:IJWF2007} and FARSITE~\cite{Finney:FARSITE} is provided in reference~\cite{Bova:IJWF2015}.

Because the spread rate of the fire front depends in part on the local wind conditions, a well resolved wind field over complex terrain should improve the accuracy of the overall model. The de facto physics-based technique for practical simulation of wind is Large Eddy Simulation (LES)~\cite{Linn:2007,coen_2013,Mell:IJWF2007}. For fire simulation over complex terrain, a range of techniques are being developed, in particular, in the coupling of the fire and the wind. For fast simulation of outdoor fires, an immersed boundary method has been interfaced with a Level Set scheme~\cite{Bova:IJWF2015} and the Rothermel fire spread model. The wind speed input for Rothermel's model results from the local velocity over the terrain which is affected by fire buoyancy and flow evolution. These are in turn also influenced by wind atmospheric conditions, accounted for within simulations using a mean wind forcing concept. This is a first step at implementing a level set approach in an LES wind calculation with combustion. It is recognized that in the empirically-based development of the Rothermel formula the wind speed is not influenced by the local buoyancy induced flow.

@@

For fire modeling in the gas phase, FDS employs a low Mach approximation for thermally driven buoyant and stratified flows~\cite{Rehm:1}. Under this assumption the pressure field $p$ can be viewed as the summation of two components: a background pressure $\bar{p}(z,t)$ used in the equation of state (ideal gas law), and a hydrodynamic pressure $\tilde{p}(x,y,z,t)$ resulting from fluid motion.

Consider a mixture composed of $N$ chemical species $\alpha$, moving on a fixed point $\mathbf{x}$ in space with a mass weighted average velocity $\mathbf{u}(\mathbf{x},t)$. If $\rho(\mathbf{x},t)$ is the mixture density, the mass fraction for species $\alpha$ is $Y_\alpha = \rho_\alpha / \rho$, where $\rho_\alpha(\mathbf{x},t)$ is the species mass density and $\rho = \sum \rho_\alpha$. The scalar transport and momentum equations take the form:
\begin{eqnarray}
   \frac{\partial \rho Y_\alpha}{ \partial t} + \nabla \cdot ( \rho Y_\alpha  \mathbf{u} ) &=& - \nabla \cdot \mathbf{J_{d}}_\alpha + \dot{m}_\alpha'''  +
    \dot{m}_{b,\alpha}'''\; , \; \alpha=1,\dots,N \label{eqn:spectran} \\
    \frac{\partial \mathbf{u}}{\partial t} - \mathbf{u} \times \boldsymbol{\omega} + \nabla H - \tilde{p} \, \nabla \left( 1/\rho\right) &=&
    \frac{1}{\rho} \left[ (\rho-\rho_0) \mathbf{g} + \mathbf{f}_{b} + \nabla \cdot \boldsymbol{\tau}^{dev} \right] \label{eqn:momtran}
\end{eqnarray}
where $\mathbf{J_{d}}_\alpha=- \rho D_\alpha \boldsymbol{\nabla} Y_\alpha$ is the diffusive flux for component $\alpha$. Fick's Law for binary diffusion with respect to a background species is assumed, and $D_\alpha$ is the diffusivity of $\alpha$ with respect to the background species.  Mass fractions $Y_\alpha$, solution of equations~\eqref{eqn:spectran}, must obey realizability constraints $0<Y_\alpha<1$ and $\sum_\alpha Y_\alpha=1$. In equations~\eqref{eqn:spectran} starting from a realizable solution, realizability implies $\sum_\alpha \mathbf{J_{d}}_\alpha = 0$. Therefore, to enforce realizability, errors in diffusive transport are lumped into the most abundant species locally~\cite{McDermott:2015}. The volumetric combustion source term for species $\alpha$ is $\dot{m}_\alpha'''(\mathbf{x},t)$  and $\dot{m}_{b,\alpha}'''$ is the contribution to species $\alpha$ from subgrid particle gasification. These correspond to the amount of mass per unit volume and time of species $\alpha$ being added or substracted in a given point $\mathbf{x}$ due to chemical reaction or gasification of solid particles (or droplets), respectively.  In the momentum equation the term $H=|\mathbf{u}|^2/2 + \tilde{p}/\rho$, $\tilde{p}$ is the perturbation pressure, and $\rho_0(z)$ is the modeled height dependent background density of the atmosphere.  Also, $\boldsymbol{\omega}$ refers to the vorticity field. The gravity vector is $\mathbf{g}=(0,0,g_z)$, the buoyancy force term is $ (\rho-\rho_0) \mathbf{g}$, $\mathbf{f}_{b}$ is the term contributed by modeled particle drag forces, and $\boldsymbol{\tau}^{dev}$ is the deviatoric stress tensor accounting for molecular and subgrid turbulent stresses parameterized by an effective eddy viscosity model, discussed below in Sec.~\ref{sec:subgrid}.

@@

Section~\ref{particle_model} discusses how Lagrangian particles are used to represent vegetation like leaves, pine needles and so on. These particles absorb and emit thermal radiation, and the underlying assumptions are described in that section.

@@

In the formal derivation of the LES equations (see, e.g, \cite{Pope:2000,FDS_Math_Guide}) the filtered nonlinear terms (advection terms, mean chemical source term, and the unresolved boundary fluxes) require subgrid-scale parameterizations.  In this section we will temporarily utilize the LES filter formalism as well as Cartesian tensor index notation to clarify the terms used in the governing equations.

@@

The eddy viscosity $\mu_t$ is an important quantity in the simulation as it affects all the turbulent transport and subgrid mixing time scales, which, as we will see below, are crucial to the combustion model.  In FDS, the eddy viscosity is modeled as \cite{Pope:2000,Deardorff:1980}
\begin{equation}
\mu_t = \bar{\rho} C_\nu \Delta \sqrt{k_{sgs}}
\end{equation}
with the constant taken as $C_\nu = 0.1$ (this value can be derived assuming production equals dissipation and a model Kolmogorov spectrum \cite{Pope:2000}; we also test the model against decaying isotropic turbulence \cite{CBC}).  In practice, the filter scale $\Delta$ is taken as the cube root of the cell volume.  Deardorff \cite{Deardorff:1980} solved a transport equation for the subgrid kinetic energy per unit mass, $k_{sgs}$, in order to account for subgrid buoyancy. In FDS, the density field is sufficiently resolved to account for buoyant plume accelerations, and therefore we find that an algebraic closure for $k_{sgs}$ is sufficient.  We utilize a scale-similarity idea, following the work of Bardina \cite{Bardina:1980}, and model the subgrid kinetic energy per unit mass as follows:
\begin{equation}
k_{sgs} = \frac{1}{2}(\tilde{u}_i - \hat{\tilde{u}}_i)(\tilde{u}_i - \hat{\tilde{u}}_i)
\end{equation}
where the hat represents a test filter at a scale $2\Delta$.  For the first Cartesian cell off the wall, where a test filter is not well defined, the eddy viscosity is taken from the WALE model \cite{Nicoud:1999}.

Subgrid advection of mass and heat are also modeled using gradient diffusion.  The turbulent diffusivities use constant turbulent Schmidt and Prandtl numbers, both set to 0.5; this simplification is justified based on high resolution simulations of fire plumes (see, e.g., \cite{Maragkos:2020}). For mass diffusion of species $\alpha$ in direction $i$ we have
\begin{equation}
J_{\alpha,i} = \bar{J}_{\alpha,i} + J_{\alpha,i}^{sgs} = -\bar{\rho}(D_\alpha + \frac{\mu_t}{\mathrm{Sc}_t}) \frac{\partial \tilde{Y}_\alpha}{\partial x_i}
\end{equation}
We use effective thermal conductivity and mass diffusivity in the heat flux, which may be written as
\begin{equation}
{\dot{q}''}_i = -(\bar{k} + \frac{\mu_t \bar{c}_p}{\mathrm{Pr}_t}) \frac{\partial \widetilde{T}}{\partial x_i} - \sum_\alpha \tilde{h}_{{\rm s},\alpha} ( \bar{\rho} D_\alpha + \frac{\mu_t}{\mathrm{Sc}_t} ) \frac{\partial \tilde{Y}_\alpha }{\partial x_i}
\end{equation}

@@

FDS treats combustion chemistry using a simplified approach.  We track only Fuel, Air, and Products, which we refer to as ``lumped species''.  Lumped species are groups of species that transport and react together and are always found in the same proportion.  For example, Air is a lumped species made up of 23 \% O$_2$ and 77 \% N$_2$ by mass.  For vegetation, a surrogate hydrocarbon Fuel is usually assumed.  For example, cellulose may be assumed to decompose to a volatile fuel gas with the formula $\C_{6} \H_{10} \O_{5}$.  The Products are defined stoichiometrically, often including prescribed yields of soot (for smoke) and carbon monoxide that may be obtained from empirical relationships for the specific fuel or from other sources (e.g., \cite{SFPE:Mulholland}).  Then, usually, a single reaction is tracked\footnote{More complicated reaction schemes are possible in the code, but for this work a single reaction is sufficient.}:
\begin{equation}
\mathrm{Fuel} + s \, \mathrm{Air} \rightarrow (1+s) \,\mathrm{Products}
\end{equation}
where $s$ is the mass stoichiometric coefficient.

@@

\subsubsection{Unresolved Boundary Fluxes}

In this section we discuss the wall functions used to close the mass, momentum, and energy fluxes on solid boundaries.  Mass transfer of moisture and fuel from the surface is discussed in more detail below in Sec.~\ref{sec:firespread} where the various flames spread models are compared (which is really the main point of this paper).

The scaled streamwise velocity is denoted $u^+ = u/u_\tau$, where the friction velocity is $u_\tau \equiv \sqrt{\tau_w/\rho}$ and $\tau_w$ is the wall stress needed to close the momentum equation, Eq.~(\ref{eqn:momtran}).  For rough walls FDS employs the log law presented in \cite{Pope:2000},
\begin{equation}
\label{eqn_roughwallloglaw}
u^+ = \frac{1}{\kappa} \ln \left(\frac{y}{s}\right) + \tilde{B}(s^+)
\end{equation}
where $\kappa = 0.41$ is the von K\'arm\'an constant, $s^+ = s/\delta_\nu$ is the roughness length in viscous units, $s$ is the dimensional ``sand grain'' roughness, and $\delta_\nu = \mu/(\rho u_\tau)$ is the viscous length scale. The distance to the wall, $y$, is taken as $\delta y/2$ for the first off-wall grid cell.  The parameter $\tilde{B}$ varies with $s^+$ but attains a constant value $B_2=8.5$ in the fully rough limit.  In FDS, we implement $\tilde{B}$ as the following piece-wise function:
\begin{equation}
\tilde{B} = \left\{ \begin{array}{lll} B + (1/\kappa)\ln(s^+) & \mbox{for} & s^+ < 5.83 \\
\tilde{B}_{\rm max} & \mbox{for} & 5.83 \le s^+ < 30.0 \\
B_2 & \mbox{for} & s^+ \ge 30.0 \end{array} \right.
\end{equation}
where $\tilde{B}_{\rm max} = 9.5$.

@@

Again, this simple approach is viable because FDS usually resolves the thermal flow fields that would necessitate added complexity.  However, in future work we plan to explore improvements to wall functions that may lead to better results at coarser grid resolution.

@@

\subsection{Scalar transport and energy discretization in unstructured region}

@@

For a control volume that does not change with time, the time derivative and source terms are approximated by

@@

  \int_{\Omega_{ii}} {\frac{\partial \rho Y_\alpha}{\partial t}} \, \dV \approx \frac{\partial \widetilde{\rho Y_\alpha }_{ii}}{\partial t} V_{ii} \quad ; \quad
  \int_{\Omega_{ii}} { \left( \dot{m}_\alpha''' +\dot{m}_{b,\alpha}''' \right)} \, \dV \approx \left(\widetilde{ \dot{m}_\alpha''' }_{ii}+ \widetilde{\dot{m}_{b,\alpha}'''}_{ii} \right) V_{ii} \label{eq:intcons}

@@

where $V_{ii}$ is the volume of cell $ii$ and tildes imply cell averages. In the following the notation is simplified by dropping tildes, observing that quantities are cell or face averaged. Consider the FV discretization of the diffusive term of Eq.~(\ref{eq:intconvdiff}) on cut-cell $ii$ of Fig.~\ref{Fig:figure_2}:

@@

where the advective flux for face $k$ is $\left( \rho Y_\alpha \mathbf{u} \right)_k = \overline{[\rho Y_\alpha]}_k \mathbf{u}_k$, and the over bar in $\overline{[\rho Y_\alpha]}_k$ means it is a flux-limited interpolation to the cut-face~\cite{FDS_Users_Guide}. In the cut-cell region a Godunov flux limited interpolation for the advective term is used. Flux limited interpolation is one of the fundamental components of stable numerical schemes for hyperbolic equations, of which mass transport inherits its mathematical properties.

@@

where the over line terms refer to flux limited interpolation of corresponding scalars, and terms defined with subscript $ii$ refer to cell defined quantities. All heat and mass fluxes and scalars are assumed known. Also, the vertical velocity $w_{ii}$ is interpolated to the cut-cell centroid. Details of the FDS time integration scheme can be found in references~\cite{FDS_Users_Guide,mcdermo_2014}. For each RK2 substep, the species transport equations are advanced in all FDS Cartesian cells, and then the solution (explicit fluxes and scalar densities) recomputed on the unstructured cut-cell region. A similar procedure is done for the thermodynamic divergence.

@@

As an explicit time integrator is used, in general there will arise cut-cells whose small-size will penalize severely the time step. These cut-cells are linked to larger surrounding cells. Cell linking in the context of momentum equations can be found in Ref.~\cite{kirk_2003}. In FDS, the momentum equations and pressure Poisson equation are solved on the Cartesian mesh. Therefore, an immersed boundary method is used to reconstruct velocities in the cut-cell region. A divergence integral equivalence argument is used to transfer the divergence from cut-cells to the underlying Cartesian cells in order to build the source term of the Poisson equation for the pressure.

\subsection{Immersed boundary method on cut-cell region and wall modeling}

@@

The previous equations are advanced in time using a fractional step method. As illustration, consider their Forward Euler (FE) update from $t_n$ to $t_{n+1}=t_n + \Delta t$. Given $ \mathbf{u}^n=\mathbf{u}(\mathbf{x},t_n)$, $\nabla \cdot \mathbf{u}^{n+1} = \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{\rm th}$ known
\begin{eqnarray}
  \frac{\mathbf{u}^{n+1}-\mathbf{u}^{n}}{\Delta t} &=& - \mathbf{F}^n -  \boldsymbol{\nabla} H^n \label{eq:LowMachMomEu}\\
  \nabla \cdot \mathbf{u}^{n+1} &=& \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{\rm th} \label{eq:LowMachDivEu}
\end{eqnarray}
where $\mathbf{u}^{n+1}$ represents a numerical solution at time $t_{n+1}$. This discrete FE update corresponds to the first sub-step of the FDS explicit RK2 integrator. As the potential field $H(\mathbf{x},t)$ does not have a time evolution equation, it is assumed responsible for enforcing the divergence condition and used on the projection step. Taking the divergence of Eq.~\eqref{eq:LowMachMomEu} and considering  the constraint, Eq.~\eqref{eq:LowMachDivEu}, the two steps of the method are
\begin{enumerate}
  \item Solve Poisson equation for $H^n$:

>>>>

\usepackage[export]{adjustbox}
\usepackage{url}

@@

Wildland fire modeling typically focuses on tracking the fire front, which can have a complex shape that moves and deforms based on local conditions like fuel loading, moisture content, wind, and terrain. The fire front requires a mathematical representation. The Lagrangian approach is based on a polygonal mesh of control markers that represents the front~\cite{Finney:FARSITE,Bova:IJWF2015}. On the other hand, an Eulerian representation allows for a fixed two-dimensional mesh on which the fire front is defined implicitly by a scalar field. This scalar field is called the level set function and numerical methods that compute the function are called Level Set Methods (LSM)~\cite{Sethian:1999,Osher:2006}. Some wildfire solvers that use LSM to track the fire front are described in Refs.~\cite{coen_2013,Bova:IJWF2015,FDS_Users_Guide,LAUTENBERGER_2013}. A practical comparison of the two techniques as implemented in the Fire Dynamics Simulator (FDS)~\cite{Mell:IJWF2007} and FARSITE~\cite{Finney:FARSITE} is provided in Ref.~\cite{Bova:IJWF2015}.

Because the spread rate of the fire front depends in part on the local wind conditions, a well resolved wind field over complex terrain should improve the accuracy of the overall model. The de facto physics-based technique for practical simulation of wind is Large Eddy Simulation (LES)~\cite{Linn:2007,coen_2013,Mell:IJWF2007}. For fire simulation over complex terrain, a range of techniques are being developed, in particular, in the coupling of the fire and the wind. For fast simulation of outdoor fires, an immersed boundary method has been interfaced with a level set scheme~\cite{Bova:IJWF2015} and the Rothermel fire spread model. The wind speed input for Rothermel's model results from the local velocity over the terrain which is affected by fire buoyancy and flow evolution. These are in turn also influenced by wind atmospheric conditions, accounted for within simulations using a mean wind forcing concept. This is a first step at implementing a level set approach in an LES wind calculation with combustion. It is recognized that in the empirically-based development of the Rothermel formula the wind speed is not influenced by the local buoyancy induced flow.

@@

For fire modeling in the gas phase, FDS employs a low Mach approximation for thermally driven buoyant and stratified flows~\cite{Rehm:1}. Under this assumption the pressure field $p$ can be viewed as the summation of two components: a background hydrostatic pressure field $\bar{p}(z,t)$ used in the equation of state (ideal gas law) and a hydrodynamic pressure $\tilde{p}(x,y,z,t)$ driving fluid motion.

Consider a mixture composed of $N$ chemical species $\alpha$, moving on a fixed point $\mathbf{x}$ in space with a mass weighted average velocity $\mathbf{u}(\mathbf{x},t)$. If $\rho(\mathbf{x},t)$ is the mixture density, the mass fraction for species $\alpha$ is $Y_\alpha = \rho_\alpha / \rho$, where $\rho_\alpha(\mathbf{x},t)$ is the species mass density and $\rho = \sum \rho_\alpha$. The scalar transport and momentum equations take the form:
\begin{eqnarray}
   \frac{\partial \rho Y_\alpha}{ \partial t} + \nabla \cdot ( \rho Y_\alpha  \mathbf{u} ) &=& - \nabla \cdot \mathbf{J_{d}}_\alpha + \dot{m}_\alpha'''  +
    \dot{m}_{b,\alpha}'''\; , \; \alpha=1,\dots,N \label{eqn:spectran} \\
    \frac{\partial \mathbf{u}}{\partial t} - \mathbf{u} \times \boldsymbol{\omega} + \nabla H - \tilde{p} \, \nabla \left( 1/\rho\right) &=&
    \frac{1}{\rho} \left[ (\rho-\rho_0) \mathbf{g} + \mathbf{f}_{b} + \nabla \cdot \boldsymbol{\tau}^{dev} \right] \label{eqn:momtran}
\end{eqnarray}
where $\mathbf{J_{d}}_\alpha=- \rho D_\alpha \boldsymbol{\nabla} Y_\alpha$ is the diffusive flux for component $\alpha$. Fick's Law for binary diffusion with respect to a background species is assumed, and $D_\alpha$ is the diffusivity of $\alpha$ with respect to the background species.  Mass fractions $Y_\alpha$, solution of equations~\eqref{eqn:spectran}, must obey realizability constraints $0<Y_\alpha<1$ and $\sum_\alpha Y_\alpha=1$. In equations~\eqref{eqn:spectran} starting from a realizable solution, realizability implies $\sum_\alpha \mathbf{J_{d}}_\alpha = 0$. Therefore, to enforce realizability, errors in diffusive transport are lumped into the most abundant species locally~\cite{McDermott:2015}. The volumetric combustion source term for species $\alpha$ is $\dot{m}_\alpha'''(\mathbf{x},t)$;  the term $\dot{m}_{b,\alpha}'''$ is the contribution to species $\alpha$ from subgrid particle gasification. These correspond to the amount of mass per unit volume and time of species $\alpha$ being added or substracted in a given point $\mathbf{x}$ due to chemical reaction or gasification of solid particles (or droplets), respectively.  Details of the combustion model are provided below in Sec.~\ref{sec:subgrid}.

In the momentum equation the term $H=|\mathbf{u}|^2/2 + \tilde{p}/\rho$, $\tilde{p}$ is the perturbation pressure, and $\rho_0(z)$ is the modeled height dependent background density of the atmosphere.  Also, $\boldsymbol{\omega}$ refers to the vorticity field. The gravity vector is $\mathbf{g}=(0,0,g_z)$, the buoyancy force term is $ (\rho-\rho_0) \mathbf{g}$, $\mathbf{f}_{b}$ is the term contributed by modeled particle drag forces, and $\boldsymbol{\tau}^{dev}$ is the deviatoric stress tensor accounting for molecular and subgrid turbulent stresses parameterized by an effective eddy viscosity model, discussed below in Sec.~\ref{sec:subgrid}.

@@

Section~\ref{particle_model} discusses how Lagrangian particles are used to represent vegetation like leaves, pine needles and other subgrid objects. These particles absorb and emit thermal radiation, and the underlying assumptions are described in that section.

@@

In the formal derivation of the LES equations (see, e.g, \cite{Pope:2000,FDS_Math_Guide}) the filtered nonlinear terms (advection terms, mean chemical source term, and the unresolved boundary fluxes) require subgrid-scale parameterizations.  In this section, LES filter notation as well as Cartesian tensor index notation are used to clarify the terms used in the governing equations.

@@

The eddy viscosity $\mu_t$ is an important quantity in the simulation as it affects all the turbulent transport and subgrid mixing time scales, which, as shown below, are crucial to the combustion model.  In FDS, the eddy viscosity is modeled as \cite{Pope:2000,Deardorff:1980}
\begin{equation}
\mu_t = \bar{\rho} C_\nu \Delta \sqrt{k_{sgs}}
\end{equation}
with the constant taken as $C_\nu = 0.1$ (this value can be derived assuming production equals dissipation and a model Kolmogorov spectrum \cite{Pope:2000}; the model is also tested against decaying isotropic turbulence \cite{CBC}).  In practice, the filter scale $\Delta$ is taken as the cube root of the cell volume.  Deardorff \cite{Deardorff:1980} solved a transport equation for the subgrid kinetic energy per unit mass, $k_{sgs}$, in order to account for subgrid buoyancy. In FDS, the density field is sufficiently resolved to account for buoyant plume accelerations, and therefore it is found that an algebraic closure for $k_{sgs}$ is sufficient.  Following the work of Bardina \cite{Bardina:1980}, a scale-similarity idea is used; the subgrid kinetic energy per unit mass is modeled as follows:
\begin{equation}
k_{sgs} = \frac{1}{2}(\tilde{u}_i - \hat{\tilde{u}}_i)(\tilde{u}_i - \hat{\tilde{u}}_i)
\end{equation}
where the hat represents a test filter at a scale $2\Delta$.  For the first Cartesian cell off the wall, where a test filter is not well defined, the eddy viscosity is taken from the WALE model \cite{Nicoud:1999}.

Subgrid advection of mass and heat are also modeled using gradient diffusion.  The turbulent diffusivities use constant turbulent Schmidt and Prandtl numbers, both set to 0.5; this simplification is justified based on high resolution simulations of fire plumes (see, e.g., \cite{Maragkos:2020}). For mass diffusion of species $\alpha$ in direction $i$ the flux is
\begin{equation}
J_{\alpha,i} = \bar{J}_{\alpha,i} + J_{\alpha,i}^{sgs} = -\bar{\rho}(D_\alpha + \frac{\mu_t}{\mathrm{Sc}_t}) \frac{\partial \tilde{Y}_\alpha}{\partial x_i}
\end{equation}
Effective thermal conductivity and mass diffusivity are used in the heat flux, which may be written as
\begin{equation}
{\dot{q}''}_i = -(\bar{k} + \frac{\mu_t \bar{c}_p}{\mathrm{Pr}_t}) \frac{\partial \widetilde{T}}{\partial x_i} - \sum_\alpha \tilde{h}_{{\rm s},\alpha} ( \bar{\rho} D_\alpha + \frac{\mu_t}{\mathrm{Sc}_t} ) \frac{\partial \tilde{Y}_\alpha }{\partial x_i}
\end{equation}

@@

FDS treats combustion chemistry using a simplified approach.  Only Fuel, Air, and Products, referred to as ``lumped species'', are tracked.  Lumped species are groups of species that transport and react together and are always found in the same proportion.  For example, Air is a lumped species made up of 23 \% O$_2$ and 77 \% N$_2$ by mass.  For vegetation, a surrogate hydrocarbon Fuel is usually assumed.  For example, cellulose may be assumed to decompose to a volatile fuel gas with the formula $\C_{6} \H_{10} \O_{5}$.  The Products are defined stoichiometrically, often including prescribed yields of soot (for smoke) and carbon monoxide that may be obtained from empirical relationships for the specific fuel or from other sources (e.g., \cite{SFPE:Mulholland}).  Then, usually, a single reaction is tracked\footnote{More complicated reaction schemes are possible in the code, but for this work a single reaction is sufficient.}:
\begin{equation}
\mathrm{Fuel} + s \, \mathrm{Air} \rightarrow (1+s) \,\mathrm{Products}
\end{equation}
where $s$ is the mass stoichiometric coefficient.

@@

\subsubsection{Unresolved Boundary Fluxes}
\label{sec:boundflx}
In this section, the wall functions used to close the mass, momentum, and energy fluxes on solid boundaries are discussed.  Mass transfer of moisture and fuel from the surface is discussed in more detail below in Sec.~\ref{sec:firespread} where the various flames spread models are compared (which is really the main point of this paper).

The scaled streamwise velocity is denoted $u^+ = u/u_\tau$, where the friction velocity is $u_\tau \equiv \sqrt{\tau_w/\rho}$ and $\tau_w$ is the wall stress needed to close the momentum equation, Eq.~(\ref{eqn:momtran}).  For rough walls FDS employs the log law presented in \cite{Pope:2000},
\begin{equation}
\label{eqn_roughwallloglaw}
u^+ = \frac{1}{\kappa} \ln \left(\frac{y}{s}\right) + \tilde{B}(s^+)
\end{equation}
where $\kappa = 0.41$ is the von K\'arm\'an constant, $s^+ = s/\delta_\nu$ is the roughness length in viscous units, $s$ is the dimensional ``sand grain'' roughness, and $\delta_\nu = \mu/(\rho u_\tau)$ is the viscous length scale. The distance to the wall, $y$, is taken as $\delta y/2$ for the first off-wall grid cell.  The parameter $\tilde{B}$ varies with $s^+$ but attains a constant value $B_2=8.5$ in the fully rough limit.  In FDS, $\tilde{B}$ is implemented as the following piece-wise function:
\begin{equation}
\tilde{B} = \left\{ \begin{array}{lll} B + (1/\kappa)\ln(s^+) & \mbox{for} & s^+ < 5.83 \\
\tilde{B}_{\rm max} & \mbox{for} & 5.83 \le s^+ < 30.0 \\
B_2 & \mbox{for} & s^+ \ge 30.0 \end{array} \right.
\end{equation}
where $\tilde{B}_{\rm max} = 9.5$.

@@

Again, this simple approach is viable because FDS usually resolves the thermal flow fields that would necessitate added complexity.  However, future work is planned to explore improvements to wall functions that may lead to better results at coarser grid resolution.

@@

\subsection{Cut-cell Scalar Transport and Energy Discretization}

@@

For a cut-cell control volume the time derivative and source terms are approximated by

@@

  \int_{\Omega_{ii}} {\frac{\partial \rho Y_\alpha}{\partial t}} \, \dV \approx \frac{\partial \overline{\rho Y_\alpha }_{ii}}{\partial t} V_{ii} \quad ; \quad
  \int_{\Omega_{ii}} { \left( \dot{m}_\alpha''' +\dot{m}_{b,\alpha}''' \right)} \, \dV \approx \left(\overline{ \dot{m}_\alpha''' }_{ii}+ \overline{\dot{m}_{b,\alpha}'''}_{ii} \right) V_{ii} \label{eq:intcons}

@@

where $V_{ii}$ is the volume of cell $ii$ and the overlines imply cell averages. In the following the notation is simplified by dropping overlines, observing that quantities are cell or face averaged. Consider the FV discretization of the diffusive term of Eq.~(\ref{eq:intconvdiff}) on cut-cell $ii$ of Fig.~\ref{Fig:figure_2}:

@@

where the advective flux for face $k$ is $\left( \rho Y_\alpha \mathbf{u} \right)_k = \overline{[\rho Y_\alpha]}_k \mathbf{u}_k$, and the over bar in $\overline{[\rho Y_\alpha]}_k$ means it is a flux-limited interpolation to the cut-face~\cite{FDS_Users_Guide}. In the cut-cell region a Godunov flux limited interpolation for the advective term is used. Flux limited interpolation is one of the fundamental components of stable numerical schemes for hyperbolic equations~\cite{leveque_2002}, of which mass transport inherits its mathematical properties.

@@

where the over line terms refer to flux limited interpolation of corresponding scalars, and terms defined with subscript $ii$ refer to cell defined quantities. All heat and mass fluxes and scalars are assumed known. Also, the vertical velocity $w_{ii}$ is interpolated to the cut-cell centroid. Details of the FDS time integration scheme can be found in references~\cite{FDS_Users_Guide,mcdermo_2014}. For each RK2 substep, the species transport equations are advanced in all FDS Cartesian cells, and then the solution (explicit fluxes and scalar densities) is recomputed on the unstructured cut-cell region. A similar procedure is done for the thermodynamic divergence.

@@

As an explicit time integrator is used, in general there will arise cut-cells whose small size will penalize severely the time step. These cut-cells are linked to larger surrounding cells. Cell linking in the context of momentum equations can be found in Ref.~\cite{kirk_2003}. In FDS, the momentum equations and pressure Poisson equation are solved on the Cartesian mesh. Therefore, an immersed boundary method is used to reconstruct velocities in the cut-cell region. A divergence integral equivalence argument is used to transfer the divergence from cut-cells to the underlying Cartesian cells in order to build the source term of the Poisson equation for the pressure.

\subsection{Immersed Boundary Method and Wall Modeling}

@@

The previous equations are advanced in time using a fractional step method. As illustration, consider their Forward Euler (FE) update from $t_n$ to $t_{n+1}=t_n + \Delta t$. Given $ \mathbf{u}^n=\mathbf{u}(\mathbf{x},t_n)$, $\left( \nabla \cdot \mathbf{u} \right)^{\mathrm{th},\,n+1}$ are known:
\begin{eqnarray}
  \frac{\mathbf{u}^{n+1}-\mathbf{u}^{n}}{\Delta t} &=& - \mathbf{F}^n -  \boldsymbol{\nabla} H^n \label{eq:LowMachMomEu}\\
  \nabla \cdot \mathbf{u}^{n+1} &=& \left( \nabla \cdot \mathbf{u} \right)^{\mathrm{th},\,n+1} \label{eq:LowMachDivEu}
\end{eqnarray}
where $\mathbf{u}^{n+1}$ represents a numerical solution at time $t_{n+1}$. This discrete FE update corresponds to the first sub-step of the FDS explicit RK2 integrator. The potential field $H(\mathbf{x},t)$ does not have a time evolution equation, it is responsible for enforcing the divergence condition and used on the projection step. Taking the divergence of Eq.~\eqref{eq:LowMachMomEu} and considering  the constraint, Eq.~\eqref{eq:LowMachDivEu}, the two steps of the method are
\begin{enumerate}
  \item Solve Poisson equation for $H^n$:

\begin{equation}
   \nabla \cdot \boldsymbol{\nabla} H^n = - \left[ \frac{\left( \nabla \cdot \mathbf{u} \right)^{\mathrm{th},\,n+1} - \nabla \cdot \mathbf{u}^{n}}{\Delta t} \right] - \nabla \cdot \mathbf{F}^n \label{it:FSPoisson}
\end{equation}
